---
title: "bat_deep_learning"
output:
 # pdf_document: default
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#####install packages
```{r packages, echo=FALSE}
pkgTest <- function(x)
{
  if (x %in% rownames(installed.packages()) == FALSE) {
    install.packages(x, dependencies= TRUE,repos="http://cran.r-project.org")    
  }
  library(x, character.only = TRUE)
}
neededPackages <- c("raster", "dplyr", "sp", "magick","reticulate"
                    )
for (package in neededPackages){pkgTest(package)}

library(keras)
#install_keras(tensorflow="default")

```

```{r}
library(keras)
# library(raster)
# library(dplyr)
library(sp)
library(magick)
library(reticulate)
```


### move bat jpgs to folders
```{r move_to_folders}

do.call(file.remove,list(list.files("all_image/train/",full.names = TRUE)))

d<-getwd()
bat_images<-list.files(path="all_image/fruit_bat")

file_counter=1

for (i in bat_images) {
  # print (i)
  bat<-image_read(paste0("all_image/fruit_bat/",i))
  image_write(image=bat,path=paste0("all_image/train/","bat",".",file_counter,".jpg"),format="jpg")
  file_counter=file_counter+1
}
#<-read_image

```

### move no_bat jpgs to folders
```{r move no_bat}
d<-getwd()
bat_images<-list.files(path="all_image/no_bat")

file_counter=1

for (i in bat_images) {
  # print (i)
  bat<-image_read(paste0("all_image/no_bat/",i))
  image_write(image=bat,path=paste0("all_image/train/","no_bat",".",file_counter,".jpg"),format="jpg")
  file_counter=file_counter+1
}

```

### assign file names to pics bats vs. no bats
```{r file_names}
original_dataset_dir<-("all_image/train")
base_dir<-("all_image/small")
dir.create(base_dir)

train_dir <- file.path(base_dir, "train")
dir.create(train_dir)
validation_dir <- file.path(base_dir, "validation")
dir.create(validation_dir)
test_dir <- file.path(base_dir, "test")
dir.create(test_dir)

train_bat_dir <- file.path(train_dir, "bat")
dir.create(train_bat_dir)

train_no_bat_dir <- file.path(train_dir, "no_bat")
dir.create(train_no_bat_dir)

validation_bat_dir <- file.path(validation_dir, "bat")
dir.create(validation_bat_dir)

validation_no_bat_dir <- file.path(validation_dir, "no_bat")
dir.create(validation_no_bat_dir)

test_bat_dir <- file.path(test_dir, "bat")
dir.create(test_bat_dir)

test_no_bat_dir <- file.path(test_dir, "no_bat")
dir.create(test_no_bat_dir)



fnames <- paste0("bat.", 1:14336, ".jpg")
file.copy(file.path(original_dataset_dir, fnames), 
          file.path(train_bat_dir),overwrite = TRUE) 

fnames <- paste0("bat.",14336:18432, ".jpg")
file.copy(file.path(original_dataset_dir, fnames), 
          file.path(validation_bat_dir),overwrite = TRUE)

fnames <- paste0("bat.", 18432:20481, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
          file.path(test_bat_dir),overwrite = TRUE)

fnames <- paste0("no_bat.", 1:13225, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
          file.path(train_no_bat_dir),overwrite = TRUE)

fnames <- paste0("no_bat.", 13225:17003, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
          file.path(validation_no_bat_dir),overwrite = TRUE) 

fnames <- paste0("no_bat.", 17003:18893, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
          file.path(test_no_bat_dir),overwrite = TRUE)

```


```{r}
original_dataset_dir<-("all_image/train")
base_dir<-("all_image/small")

train_dir <- file.path(base_dir, "train")
validation_dir <- file.path(base_dir, "validation")
test_dir <- file.path(base_dir, "test")

train_bat_dir <- file.path(train_dir, "bat")

train_no_bat_dir <- file.path(train_dir, "no_bat")

validation_bat_dir <- file.path(validation_dir, "bat")

validation_no_bat_dir <- file.path(validation_dir, "no_bat")

test_bat_dir <- file.path(test_dir, "bat")

test_no_bat_dir <- file.path(test_dir, "no_bat")

```

##read in vgg16 -- https://jjallaire.github.io/deep-learning-with-r-notebooks/notebooks/5.3-using-a-pretrained-convnet.nb.htmls
```{r vgg}
library(keras)
conv_base <- application_vgg16(
  weights = "imagenet",
  include_top = FALSE,
  input_shape = c(150, 150, 3)#comment this out so that we can use images of different size
)
```
##unfreeze weights from block 3 so that rest of conv_base can be fine-tuned
```{r unfreeze}
unfreeze_weights(conv_base, from = "block3_conv1")
```

##add dense layers to conv_base
```{r dense}
model <- keras_model_sequential() %>% 
  conv_base %>% 
  layer_flatten() %>% 
  layer_dense(units = 256, activation = "relu") %>% 
  #layer_batch_normalization()%>%
    #layer_dropout(0.5) %>%

  layer_dense(units = 1, activation = "sigmoid")
```

##make image_data_generator that does image augmentation
```{r augment}
train_datagen = image_data_generator(
  rescale = 1/255,
  rotation_range = 40,
  width_shift_range = 0.2,
  height_shift_range = 0.2,
  shear_range = 0.2,
  zoom_range = 0.2,
  horizontal_flip = TRUE,
  fill_mode = "nearest"
  
)


test_datagen <- image_data_generator(rescale = 1)
train_generator <- flow_images_from_directory(
  train_dir,
  train_datagen,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)
validation_generator <- flow_images_from_directory(
  validation_dir,
  test_datagen,
  target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)
```

##fit the model
```{r fit}
model %>% compile(
  loss = "binary_crossentropy",
  optimizer = optimizer_rmsprop(lr = 1e-5),
  metrics = c("accuracy")
)
history <- model %>% fit_generator(
  train_generator,
  steps_per_epoch = 100,
  epochs =20,
  validation_data = validation_generator,
  validation_steps = 50
)
```

##save the model
```{r save}
save_model_hdf5(model,"deep_learning_bats.h5")
```

##plot accuracy and loss for training and validation
```{r plot_accuracy}
plot(history)
```

##load the model
```{r load}
model<-load_model_hdf5("deep_learning_bats.h5", custom_objects = NULL, compile = TRUE)
```
##rename files in random_bat_pictures folder
```{r rename_test}
num_files = length(list.files("all_image/random_bat_pictures/folder"))
print(num_files)
file_rename_list = paste0(seq(from = 1, to = num_files), ".jpg")
subdirName = "all_image/random_bat_pictures/folder"
oldNames = list.files(subdirName)
file.rename(from = file.path(subdirName, oldNames), to = file.path(subdirName, file_rename_list))

#file.rename(list.files("all_image/random_bat_pictures/folder"),  file_rename_list)
list.files("all_image/random_bat_pictures/folder")
```


##make predictions in R
```{r predictions}
train_datagen = image_data_generator(
  rescale = 1/255,
  rotation_range = 40,
  width_shift_range = 0.2,
  height_shift_range = 0.2,
  shear_range = 0.2,
  zoom_range = 0.2,
  horizontal_flip = TRUE,
  fill_mode = "nearest"
)

#train_datagen = image_data_generator(
#  rescale = 1/255
#)


#get data from https://www.kaggle.com/c/dogs-vs-cats/data. Put eight image into folder test1image
test_dir<-"all_image/random_bat_pictures"

test_generator <- flow_images_from_directory(
  test_dir,                  # Target directory  
  train_datagen,              # Data generator
  target_size = c(150, 150),  # Resizes all images to 150 × 150
  batch_size = 1,
  class_mode = "binary",
  shuffle = FALSE# binary_crossentropy loss for binary labels
)

num_test_images = 29
y <- predict_generator(model, test_generator, steps=num_test_images, 
                       verbose =1)
y
summary(y)

```

##view all the images with values close to 0 
```{r}
indicesNoBatPrediction = which(y < 0.5)
subdirName = "all_image/random_bat_pictures/folder/"

fileNames = list.files("all_image/random_bat_pictures/folder/")
for (image in 1:length(indicesNoBatPrediction) ){
  fileNameOne = paste0(subdirName,
                     fileNames[indicesNoBatPrediction[image]])
  print(fileNameOne)
  r <- raster(fileNameOne)
  plot(r, col = gray.colors(10, start = 0.3, end = 0.9, gamma = 2.2, alpha = NULL))
}
#probas_to_classes(y) function not found

```


###now plot images with values close to 1
```{r images_close_to_1}
##view all the images with values close to 0 
indicesPrediction1 = which(y >= 0.5)
subdirName = "all_image/random_bat_pictures/folder/"

fileNames = list.files("all_image/random_bat_pictures/folder/")
for (image in 1:length(indicesPrediction1) ){
  fileNameOne = paste0(subdirName,
                     fileNames[indicesPrediction1[image]])
  print(fileNameOne)
  r <- raster(fileNameOne)
  plot(r, col = gray.colors(10, start = 0.3, end = 0.9, gamma = 2.2, alpha = NULL))
}

```


##make predictions in Python
###first import keras -- nope, doesn't work to import keras
```{r}
py_module_available(keras)
py_module_available(numpy)
#keras = import(keras)
```

